\section{Introduction}
\label{sec:introduction}

Large language models are increasingly deployed as autonomous agents that call
external tools and APIs~\citep{schick2024toolformer, yao2023react}. Several
orchestration frameworks structure this behavior as explicit \emph{workflow
graphs}: LangGraph~\citep{langgraph_docs}, CrewAI~\citep{crewai_docs},
AutoGen~\citep{wu2023autogen}, and Google ADK~\citep{google_adk_docs} each
represent computation steps as nodes and allowable transitions as edges. Yet
safety enforcement in these systems remains almost entirely a \emph{runtime}
concern: tools such as NeMo Guardrails~\citep{nemo_guardrails} and
LlamaGuard~\citep{llamaguard} filter individual calls at execution time, detecting
a violation only when the offending path is exercised.

Some defects, however, are topological. Consider an email-triage agent whose
developer adds a \texttt{draft\_response} node but forgets to connect it to
\texttt{send}: normal-priority emails silently halt at a dead end. A content-level
runtime guardrail cannot catch this unless that path runs, but a static analysis
of the graph flags the dead end and emits a witness trace. In principle classical
model checkers~\citep{clarke1999modelchecking, holzmann1997spin,
cimatti2002nusmv} verify such properties, but they require hand-translating the
system into Promela or SMV---prohibitive for developers iterating on workflows.
For conventional static analyzers, decades of deployment experience show that
what decides adoption is not detection power but the rate of genuine, actionable
findings~\citep{bessey2010coverity, johnson2013static, sadowski2018lessons}. No
analogous measurement exists for agent workflows.

We build \textsc{Agentproof} to make that measurement. It extracts a unified
graph model from four frameworks' APIs with no per-workflow manual modeling,
runs six structural checks with witness traces, and compiles a temporal policy
DSL with LTL$_f$ (finite-trace) semantics~\citep{degiacomo2013ltlf} to DFAs
checked statically via a graph\,$\times$\,DFA product and enforced at runtime.
We then use it as a measurement instrument on 922 workflow graphs mined from
252 public GitHub repositories, with a stratified 119-workflow sample validated
against independently reconstructed ground-truth graphs and adversarial triage.

Because a lossy extractor both \emph{manufactures} false flags and \emph{hides}
true defects, we report two estimands separately: the \emph{precision} of the
instrument's flags (what fraction of raised flags are genuine) and the
\emph{prevalence} of defects per workflow, estimated by re-running every check
on the reconstructed ground-truth graphs so that defects the extractor missed
are counted. The distinction matters: flag precision is 1.1\% (2 of 186), but
prevalence is roughly twice what the extractor sees---4 of 119 workflows
(3.4\%)---because two genuine defects, including a payment-adjacent
missing-human-gate, were invisible to the lossy extraction. Both are small;
neither is the ${\approx}0$ a na\"ive reading of flag counts would suggest, and
the defects concentrate exactly where they matter: in application-like
workflows binding side-effecting tools (3 of 39, versus 1 of 80 among
tutorials, demos, and tests).

\paragraph{Contributions.}
\begin{itemize}\itemsep2pt
  \item \textbf{A base-rate study of topology defects in public agent
  workflows}, with the estimands kept honest: flag precision \emph{and}
  workflow-level prevalence including false negatives, measured against
  independent graph reconstructions, with repository-clustered uncertainty and
  a source-level census of the corpus (one third application-like, two thirds
  tutorials/demos/tests; Section~\ref{sec:realworld}).
  \item \textbf{\textsc{Agentproof}}, the measurement instrument: automatic
  extraction of a unified graph model from four heterogeneous frameworks, six
  structural checks with witness traces, and a seven-form policy DSL with
  precise LTL$_f$ semantics whose static product covers bad prefixes,
  end-of-trace obligations, and divergent (never-satisfiable) obligations
  (Section~\ref{sec:system}).
  \item \textbf{A quantified diagnosis}: extraction fidelity, not the check
  algorithms, is the binding constraint. Edge recall is 0.65 on real code
  (0.38 on ADK) versus near-perfect on stubs; every structural flag on
  extracted graphs is an extraction artifact, and structural flags arise only
  for the one extractor that does not impose connected topologies---so
  extractor-level structural conclusions are meaningful for LangGraph only
  (Section~\ref{sec:realworld}).
  \item \textbf{A defect-independent job} for the static machinery, stated
  with its premise: given a graph that over-approximates runtime traces, the
  product proves which policy monitors can never fire, so they can be soundly
  pruned. On the curated corpus 251 of 270 monitor instances (93\%) are
  provably inert---though on today's small workflows every proven case is
  alphabet-level, so the product's value there is soundness on branching
  topologies, not extra pruning (Section~\ref{sec:results}).
\end{itemize}
