\section{Base-Rate Study on Public GitHub Workflows}
\label{sec:realworld}

To estimate defect base rates on independently authored workflows---and to
measure extractor fidelity on code we did not write---we mined public GitHub
repositories across all four frameworks. We state the population precisely:
\emph{public GitHub files matching framework-specific code searches}, not
production deployments; we census what that population contains below.

\paragraph{Mining and corpus census.}
Using GitHub code search (via the \texttt{gh} CLI, February--March 2026, with
per-framework query sets and per-repository caps) we streamed each candidate
repository through clone, extraction, and cleanup. After removing one
repository of our own test fixtures (8 files; a contamination we detected
during revision and exclude from every number), the corpus is 922 workflow
graphs from 252 distinct repositories (LangGraph~400, AutoGen~359, CrewAI~113,
ADK~50), with provenance recorded as repository, commit SHA, and file path.
Mined workflows are small---median 2 non-sentinel nodes (mean~3.9)---and
predominantly static linear pipelines. The corpus is \emph{not} mostly
applications: 32\% of files carry tutorial/test/demo-style names, 48 topology
groups are byte-identical duplicates across repositories (170 redundant
workflows, retained but reported), and a source-level classification of the
validated sample (below) found 44\% tutorial or course material, 17\% toy
demos, 7\% test files, and 33\% application-like code. We therefore report
prevalence both pooled and stratified by this classification.

\paragraph{Validation sample and ground truth.}
On a stratified 119-workflow sample from 87 repositories (40 LangGraph, 20
CrewAI, 35 AutoGen, 24 ADK) we ran two independent LLM-agent passes
(Claude Opus~4.8 orchestrated by a multi-agent harness; all outputs released).
First, an agent reconstructed a ground-truth graph from source alone (blind to
the extractor's output, following its node-id conventions), with a
self-reported confidence (77 high, 33 medium, 9 low); we score node/edge
precision--recall and node-kind accuracy against these reconstructions.
Second, every flagged defect was triaged from source
(\textsc{real}/\textsc{extraction-artifact}/\textsc{intentional}) and an
adversarial verifier attempted to overturn each label (4.6\% overturned;
unresolved disagreements recorded as \textsc{arguable}). We acknowledge the
limitation squarely: the reference graphs are LLM reconstructions, not human
annotations, and both passes share a model family. Three mitigations bound the
risk: reconstruction confidence is reported and results are stable on the
high-confidence subset (edge recall 0.67 vs.\ 0.65 overall); every triage label
carries a source-line-cited justification that survived an independent
adversarial pass; and the label pipeline was re-audited during revision (one
label overturned of 16 re-examined). Human annotation of a subsample is the
clear next step. We report Wilson 95\% CIs for proportions and
repository-clustered bootstrap CIs (10{,}000 resamples) for means, since
workflows from one repository are not independent.

\begin{table}[t]
\centering\small
\caption{AST-extractor fidelity on real workflows ($n{=}119$), with
repository-clustered bootstrap 95\% CIs on the overall means. Perfect on
stubs; on real code fidelity is lower and strongly framework-dependent.}
\label{tab:realworld_fidelity}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{Framework} & \textbf{$n$} & \textbf{Edge P} & \textbf{Edge R} & \textbf{Kind acc} \\
\midrule
LangGraph & 40 & 0.956 & 0.682 & 0.647 \\
CrewAI    & 20 & 0.700 & 0.692 & 0.929 \\
AutoGen   & 35 & 0.957 & 0.770 & 0.900 \\
ADK       & 24 & 0.420 & 0.382 & 0.940 \\
\midrule
\textbf{Overall} & \textbf{119} & \textbf{0.805} & \textbf{0.649} & \textbf{0.829} \\
 & & [0.72,\,0.89] & [0.56,\,0.73] & [0.78,\,0.87] \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Extraction fidelity is the bottleneck.}
Node detection stays strong (overall $P{=}0.91$, $R{=}0.85$), but edge recall
falls to 0.649 and node-kind accuracy to 0.829, ranging from 0.38 edge recall
on ADK (cluster CI $[0.14,0.62]$) to 0.77 on AutoGen
(Table~\ref{tab:realworld_fidelity}). The frameworks that encode topology
\emph{implicitly}---ADK's nested \texttt{SequentialAgent}/
\texttt{ParallelAgent} composition, AutoGen's participant lists with dynamic
speaker selection---are hardest: the extractor must guess an edge set the
source never states, and for ADK it is wrong more often than right. On
LangGraph, whose edges are explicit, the dominant error is instead node-kind
(\texttt{tool}$\rightarrow$\texttt{llm}: nodes invoking tools in their body
with no declared binding). This 0.65 is a property of the static AST fallback
used at scale, not of the runtime extractors, which read the compiled
framework object: on trusted runnable examples the fallback recovers 0.71 of
the runtime extractor's LangGraph edges (matching its 0.68 on mined code),
and for ADK the runtime graph is far richer (16 vs.\ ${\sim}5$ edges).
Realizing that ceiling at scale would require sandboxed execution of untrusted
repository code, which we do not attempt.

\begin{table}[t]
\centering\small
\caption{Adversarial triage of the 186 flags the AST-extracted study raised.
Every structural flag is an extraction artifact; the blunt human-presence flag
is predominantly intentional. These rows measure flag \emph{precision}, not
defect prevalence.}
\label{tab:realworld_triage}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{Check} & \textbf{Real} & \textbf{Artifact} & \textbf{Intent.} & \textbf{Arg.} \\
\midrule
exit\_reachability    & 0 & 14 & 0  & 0 \\
reverse\_reachability & 0 & 25 & 0  & 0 \\
dead\_ends            & 0 & 24 & 0  & 1 \\
router\_shape         & 0 & 3  & 0  & 0 \\
tool\_declarations    & 0 & 5  & 0  & 0 \\
human\_presence       & 2 & 8  & 93 & 10 \\
human\_gate\_coverage & 0 & 0  & 0  & 1 \\
\midrule
\textbf{Total}        & \textbf{2} & \textbf{79} & \textbf{93} & \textbf{12} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Estimand 1: flag precision.}
Of the 186 flags raised on extracted graphs, 2 (1.1\%, Wilson 95\% CI
$[0.3,3.8]\%$) are genuine---both missing human gates on sensitive
workflows---while 42\% are extraction artifacts and 50\% intentional design
(Table~\ref{tab:realworld_triage}). \emph{Every one of the 72 structural flags
is an extraction artifact or arguable} (0 genuine; CI $[0,5.1]\%$). Structural
flags are also entirely a LangGraph phenomenon: they fire on 78\% of LangGraph
workflows but 0\% of AutoGen, CrewAI, and ADK---and this is partly
\emph{by construction}, since those three extractors emit connected topologies
and cannot produce a dead end or unreachable exit. Extractor-level structural
results are therefore evidence about LangGraph only; for the other frameworks
the structural checks are vacuous on extracted graphs, which is exactly why
the next estimand is measured on reconstructions instead.

\paragraph{Estimand 2: prevalence, including what the extractor misses.}
Flag triage cannot see false negatives: a defect the lossy extractor never
surfaces is simply absent from Table~\ref{tab:realworld_triage}. We therefore
re-ran every check on the 119 reconstructed ground-truth graphs---which can
contain dead ends and unreachable exits for all four frameworks---and
adversarially triaged all 16 resulting flags (labels: genuine, intentional, or
reconstruction error). Ten are intentional (e.g.\ deliberate wait-for-input
halt nodes in multi-turn chat loops), four are reconstruction errors, and
\textbf{two are genuine defects the extracted study missed}: a router
mis-wiring in an ADK SQL agent, and a chatbot whose SQL-executing tool is
reachable with no human gate. Combining confirmed defects from both passes
(the two originals were re-confirmed by a second adversarial pass) gives the
prevalence estimates of Table~\ref{tab:prevalence}: 4 of 119 workflows
(3.4\%), 4 of 87 repositories (4.6\%)---about \emph{twice} what flag triage
alone shows, quantifying the false-negative rate of the lossy instrument. All
four defective workflows bind side-effecting tools, and three of four are
application-like: prevalence is 7.7\% among application-like workflows versus
1.3\% among tutorials, demos, and tests. Topology defects are rare in this
population, but not artifact-of-the-instrument rare, and they concentrate
where the stakes are.

\begin{table}[t]
\centering\small
\caption{Defect prevalence (workflows with $\geq$1 genuine defect), Wilson
95\% CIs. The extractor-visible row is what flag triage alone supports; the
ground-truth row adds defects found only on reconstructed graphs.}
\label{tab:prevalence}
\begin{tabular}{@{}lrr@{}}
\toprule
\textbf{Estimate} & \textbf{Rate} & \textbf{95\% CI} \\
\midrule
Extractor-visible, per workflow      & 2/119 (1.7\%) & $[0.5, 5.9]\%$ \\
+ ground-truth graphs, per workflow  & 4/119 (3.4\%) & $[1.3, 8.3]\%$ \\
\quad per repository                 & 4/87 (4.6\%)  & $[1.8, 11.2]\%$ \\
\quad application-like only          & 3/39 (7.7\%)  & $[2.7, 20.3]\%$ \\
\quad tutorials/demos/tests          & 1/80 (1.3\%)  & $[0.2, 6.8]\%$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Risk-aware gating.}
The blunt human-presence check fires on 91\% of mined workflows with 2 of 113
firings genuine---uninformative as a detector, which is why we evaluate the
risk-aware \emph{human-gate coverage} check: flag only when a sensitive tool
(destructive/write/exec/communication/financial, by a keyword lexicon over
bound tool names) is reachable without a \textsc{human} gate. On the 922
mined graphs it fires on zero---but this is a statement about extracted
\emph{declarations}, not about the code: no mined graph declares a sensitive
binding, while source-level classification finds side-effecting tools in 32 of
119 sampled workflows, invisible to the extractor through the
\texttt{tool}$\rightarrow$\texttt{llm} kind error. On the reconstructed
graphs, where bindings are recovered from source, the same check fires three
times and one firing is the genuine SQL-gate defect above. The check's logic
is sound; its recall is bounded by extraction fidelity, which is the paper's
central diagnosis.
