\section{Controlled Evaluation and Monitor Selection}
\label{sec:results}
\label{sec:pruning}

\paragraph{Controlled detection test.}
A curated corpus of 18 workflows spanning all four frameworks, seeded with known
anti-patterns, confirms the checks fire when defects are genuinely present:
Agentproof detects every injected structural defect and, with a risk-aware
sensitive-tool set, flags 8 workflows whose sensitive path (e.g.
\texttt{account\_create}, \texttt{publish}, \texttt{auto\_remediate}) is reachable
without a human gate, versus 10 for the blunt human-presence check---declining to
flag two workflows that lack a human node but perform no sensitive action. This
corpus is a controlled test that the algorithms work, not evidence of prevalence.
Verification is sub-second even for synthetic graphs of 5{,}000 nodes (far larger
than any observed workflow); details, per-framework extractor accuracy on stubs,
and a Promela/SMV modeling-effort comparison are in the supplementary material.

\paragraph{Monitor selection via static pruning.}
The graph\,$\times$\,DFA product gives the static machinery a job that does not
depend on defects existing: deciding which runtime monitors to deploy. The
soundness premise of Section~\ref{sec:system} is required and we restrict the
claim accordingly---pruning applies to graphs whose paths over-approximate
runtime traces (here, the curated corpus; in deployment, runtime-extracted
graphs), \emph{never} to lossy AST-extracted graphs, whose missing edges could
make a live monitor look inert. Under that premise, a policy whose product
reaches no violation---by bad prefix, by termination, or by divergent
obligation---can never fire at runtime on that workflow, so its monitor need
not be deployed. Multi-tool nodes are first expanded into a complete digraph
of single-tool nodes (admitting every invocation order and subset), so the
expansion over-approximates rather than fixing one arbitrary tool order.

Across the curated corpus and 15 policies, 251 of 270 monitor instances
(93.0\%) are provably inert---a mean of 13.9 of 15 monitors skipped per
workflow, leaving runtime enforcement to evaluate only the remaining 7\%.

We report two honest limits. First, on today's small, mostly linear workflows
\emph{every} proven-inert case is alphabet-level (the policy's atoms never
appear in the workflow); under the corrected LTL$_f$ semantics the product
proves no additional case on this corpus---the single case the reachability
product ``won'' under our earlier, incorrect semantics was in fact a live
termination-obligation monitor. The product's value over a plain symbol-set
check is therefore soundness, not extra pruning: on branching and looping
topologies a symbol-set check cannot certify inertness while the product can,
and the product is what licenses the guarantee at all. Second, the fixed
policy vocabulary does not match real code---none of the 922 mined workflows
invokes any of the 15 policies' tools---so measuring pruning on mined
workflows requires instantiating policy templates against each workflow's own
tools, which we leave to future work. Static monitor selection is thus a
sound, cheap optimization whose measured leverage today is alphabet-level, and
it is the concrete bridge between pre-deployment analysis and leaner runtime
enforcement.
