\section{Related Work}
\label{sec:related}

\paragraph{Runtime agent safety.}
NeMo Guardrails~\citep{nemo_guardrails}, Guardrails AI~\citep{guardrails_ai},
LlamaGuard~\citep{llamaguard}, and Rebuff~\citep{rebuff} enforce safety at the
point of LLM output or tool call, catching content-level violations only when the
offending path runs. Sandboxed and adversarial evaluations such as
ToolEmu~\citep{ruan2024toolemu} and AgentDojo~\citep{debenedetti2024agentdojo}
measure agent risk by executing behaviors. Agentproof is complementary to both:
it analyzes structural properties of the topology before deployment, and its
monitor-selection layer (Section~\ref{sec:pruning}) decides which runtime
monitors are even needed.

\paragraph{Model checking and workflow soundness.}
Classical model checkers---SPIN~\citep{holzmann1997spin},
NuSMV~\citep{cimatti2002nusmv}, CBMC~\citep{cbmc}---verify the same class of
properties but require manual translation into Promela/SMV/annotated~C;
Agentproof extracts the model automatically. Our policy semantics follow LTL
on finite traces~\citep{degiacomo2013ltlf}, with monitoring in the style of
JavaMOP~\citep{javamop} and three-valued LTL runtime
verification~\citep{bauer2011ltl3}. Business-process soundness for Petri
nets~\citep{vanderaalst2011bpm} and BPMN~\citep{dijkman2008bpmn} studies
analogous reachability properties, but targets visual notations rather than the
programmatic API objects agent frameworks expose.

\paragraph{Base rates for static analysis.}
Industrial experience with conventional static analyzers established that
false-positive and actionability rates, not detector cleverness, decide
adoption: Coverity's deployment history~\citep{bessey2010coverity}, developer
studies of why warnings are ignored~\citep{johnson2013static}, and Google's
Tricorder lessons~\citep{sadowski2018lessons}. Our study is the agent-workflow
instance of that genre---and adds a failure mode those studies did not face:
for agent frameworks the \emph{model extraction} step itself dominates the
error budget, manufacturing 42\% of flags and hiding half the genuine defects.

\paragraph{Agent safety and evaluation.}
Work on agent safety emphasizes alignment, capability evaluation, and
governance~\citep{anthropic_rsp, metr, gabriel2024ethics}. Our contribution is
orthogonal and empirical: base-rate measurements of a specific, checkable
defect class in public agent workflows, with the estimands and their
instrument error separated.
