\section{Conclusion}
\label{sec:conclusion}

Agent frameworks encode behavior as explicit workflow graphs, and
\textsc{Agentproof} turns that structure into an analyzable, cross-framework
model with structural checks, witness traces, and a policy DSL with precise
LTL$_f$ semantics evaluated both statically (via a graph\,$\times$\,DFA
product covering bad prefixes, termination obligations, and divergent
obligations) and at runtime. Used as a measurement instrument on 922 workflows
from 252 public GitHub repositories, it yields a finding more textured than a
simple null: flags are almost never defects (precision 1.1\%), yet the
instrument also \emph{hides} defects---measured prevalence doubles, to 3.4\%
of workflows, when checks run on faithful reconstructions---and the genuine
defects concentrate in application-like workflows that bind side-effecting
tools (7.7\%, versus 1.3\% among the tutorials and demos that dominate public
GitHub). Curated-benchmark defect rates do not stand in for prevalence, and
flag counts do not stand in for either.

The diagnosis is consistent across every cut: extraction fidelity, not the
check algorithms, is the binding constraint. Edge recall is 0.65 on real code
(0.38 on implicit-topology ADK); every structural flag on extracted graphs is
an extraction artifact; three of four extractors cannot even express the
defects the structural checks target; and the risk-aware gate misses real
sensitive paths because tool bindings hide in node bodies. The static
machinery keeps a defect-independent job---under an explicit trace-containment
premise, the product soundly prunes provably-inert runtime monitors (93\% on
the curated corpus, all of it alphabet-level today)---positioning
pre-deployment analysis as a monitor-selection layer within a
static-plus-runtime safety stack.

\paragraph{Limitations.}
The population is public GitHub, not production systems; two thirds of it is
tutorial-grade, which we census and stratify but cannot fully correct.
Ground-truth graphs and triage labels are LLM reconstructions with adversarial
verification, not human annotations; confidence-stratified sensitivity checks
and released per-label justifications mitigate but do not eliminate this, and
human annotation of a subsample is the immediate next step. Prevalence CIs are
wide at $n{=}119$; the estimates are best read as ``rare, but real and
concentrated'', not as precise rates. We read these results as a
redirection---toward higher-fidelity extraction and per-workflow policy
instantiation, the direction we pursue in follow-up work on runtime shields.

\paragraph{Artifact.}
The tool, corpora, mining and validation pipeline, ground-truth
reconstructions, triage labels with justifications, and all evaluation
scripts will be released under the MIT license.

\section*{Ethical Statement}
The study mines only \emph{public} GitHub repositories, records exact provenance
(repository at commit SHA plus file path), and redistributes only the derived
graph abstractions and analysis, never third-party source. No personal data is
processed. The four genuine defects we identify are reported without
identifying maintainers in the paper; we will disclose them to the affected
repositories before publication. The intended impact is defensive---helping
developers catch topology-level safety defects before deployment and reason
about which runtime safeguards a workflow needs; the honest finding that such
defects are rare in public code, yet concentrated in side-effecting
application-like workflows, should calibrate, not inflate, claims about static
verification's role in agent safety.
