\section{Evaluation}
\label{sec:evaluation}

We evaluate Agentproof along two axes:
(i) graph extraction and structural statistics for representative agent
workflows, and (ii) microbenchmarks for temporal monitor compilation and event
evaluation throughput.

\paragraph{Case studies.}
We use three representative workflows built using real framework APIs:
(i) a LangGraph incident response triage and mitigation graph,
(ii) a hierarchical Google ADK compliance document pipeline (sequential +
parallel + loop) with a human signoff step, and
(iii) an AutoGen AgentChat change-control board topology with a human approver.
These workflows are constructed without executing any external API calls.

\paragraph{Methodology.}
For each workflow, we extract an \texttt{AgentGraph} and compute node/edge counts
as well as counts by node and edge kind. We measure extractor runtime as the
median over 200 runs (after 20 warm-up runs) on the same in-memory workflow
object. For temporal monitors, we measure compilation time as the median over
500 runs per DSL string and evaluation throughput (events/s) over a fixed
synthetic 1{,}000-event stream. We also run Agentproof's structural checks on
each extracted graph and evaluate a small temporal policy suite over three
representative traces per workflow (one safe, two unsafe).

\input{generated/case_study_table.tex}
\input{generated/case_study_verification_table.tex}
\input{generated/bench_table.tex}

\paragraph{LangGraph incident response.}
The incident response workflow uses a conditional router to branch into either
automatic mitigation or a human approval step before executing mitigation tools.
We verify \texttt{G !tool:drop\_database} and
\texttt{decision:mitigate -> F action:human\_approval}. The happy-path trace
passes, while unsafe traces halt on the forbidden tool and escalate on missing
approval between repeated mitigation decisions (Table~\ref{tab:case-study-verification}).

\paragraph{ADK compliance pipeline.}
The ADK pipeline composes sequential, parallel, and loop agents for document
collection, analysis, remediation, and human signoff. We verify
\texttt{G !tool:exfiltrate\_data} and \texttt{read\_only U action:human\_signoff}.
The monitors halt on exfiltration attempts and block writes prior to signoff.

\paragraph{AutoGen change-control board.}
The AutoGen case study models an acyclic speaker transition topology ending in
a \texttt{UserProxyAgent} approver. We verify \texttt{G !tool:send\_email} and
\texttt{decision:deploy -> F action:human\_approval}; unsafe traces halt on email
attempts and on repeated deploy decisions without a human approval event.

\paragraph{Discussion.}
Across case studies, extracted graphs are small enough that structural checks
complete in time linear in graph size. Temporal monitoring is also lightweight:
rules compile into constant-size automata and evaluation performs a constant
number of table lookups per event per rule.
