\section{Limitations and future work}
\label{sec:limitations}

\paragraph{Workflow structure vs.\ LLM semantics.}
The proposed approach verifies properties of the workflow \emph{topology} and
the event stream emitted by execution. It does not prove semantic
properties of LLM outputs (e.g., factuality or intent), which depend
on non-deterministic model behavior and prompt/context choices.

\paragraph{Extractor heuristics.}
Frameworks evolve quickly and expose different internal representations
across versions. The extractors employ heuristics (e.g., recognizing
human review nodes by naming conventions) that may not generalize to
all codebases. Section~\ref{sec:accuracy} discusses classification
accuracy in detail.

\paragraph{Runtime graph mutation (T3).}
Some frameworks allow dynamic modification of the workflow graph during
execution. If the post-mutation graph is not re-verified, static
guarantees no longer hold. Integrating re-verification hooks into
framework event systems is a natural extension.

\paragraph{Temporal DSL expressiveness.}
The DSL covers seven expression forms targeting the safety fragment of
LTL but does not support full LTL, nested temporal operators beyond
two levels, or past-time modalities. In the evaluation, all 15
concrete policies fit within the seven-form fragment (100\% coverage),
suggesting the fragment is sufficient for common agent safety
requirements. Extending to richer fragments and to full graph
$\times$ automaton product construction is future work.

\paragraph{False positive management.}
Some flagged defects may be intentional design choices (e.g., dead-end
nodes used as intentional error-halting states, or workflows that
intentionally omit human gates for low-risk tasks). Future work
includes annotation-based suppression (e.g.,
\texttt{\# agentproof: ignore dead-end} comments), severity tiers,
and configurable verification profiles. The library already supports
a \texttt{suppressions} parameter for node-level exclusions.

\paragraph{Corpus scale.}
The evaluation uses 18 curated workflows. Agentproof provides a GitHub mining
pipeline (\texttt{scripts/\allowbreak scrape\_workflows.py}) and an AST-based
fallback extractor (\texttt{scripts/\allowbreak ast\_extractor.py}) for building
larger real-world corpora. A large-scale study across hundreds of
open-source repositories would strengthen the empirical findings
and provide a reusable benchmark for the community.

