\section{Conclusion}
\label{sec:conclusion}

Agent frameworks already encode behavior as explicit workflow graphs.
Agentproof leverages this structure to enable pre-deployment verification of
safety properties without adding runtime overhead.

Agentproof provides a unified graph model with six structural
checks and witness traces, a temporal policy DSL covering the safety
fragment of LTL with both static and runtime evaluation modes, and
automatic extractors for four major agent frameworks---eliminating the
manual modeling effort required by general-purpose model checkers. In 18 curated workflows, 27\% contain
structural defects (dead ends, unreachable exits) and 55\% violate a
human-gate policy when required. All 15 temporal safety policies
evaluated fit within the seven-form DSL fragment, with verification
completing in sub-second time even for graphs of 5{,}000 nodes.

The primary technical contribution is not the graph algorithms
themselves---which are standard---but rather: (i)~the identification of
a practical new domain where these algorithms apply directly,
(ii)~the engineering of extractors that normalize four heterogeneous
framework APIs into a single analyzable representation, and (iii)~the
empirical finding that real-world agent workflow patterns contain
structural defects detectable by these methods.

Static verification does not replace runtime guardrails; the two
approaches are complementary. By catching topology-level defects
exhaustively before deployment, static analysis reduces the safety
surface that runtime enforcement must cover, making both layers more
effective.

\paragraph{Artifact availability.}
The Agentproof tool, curated corpus, temporal policies, and
all evaluation scripts are available at
\url{https://github.com/NordicAgents/AgentProof} under the MIT
license. A reproducibility script
(\texttt{scripts/reproduce\_all.sh}) runs the complete evaluation
pipeline.

