\section{Background}
\label{sec:background}

\subsection{Agent workflow graphs}
Many agent frameworks represent behavior as a directed graph, sometimes
called a \emph{state graph} or \emph{workflow graph}. Nodes correspond
to computation steps: an LLM call, a tool invocation, a
router/selector that branches based on state, or a human approval gate.
Edges correspond to transitions between steps, including conditional
branches, parallel fan-outs, and loop back-edges.

This explicit control-flow model is a key difference between
graph-orchestrated agents and monolithic ``single prompt''
applications: the graph structure can be analyzed statically without
executing the agent.

\subsection{Structural verification}
Structural properties depend only on graph topology and type
annotations. Typical examples include:
(i)~reachability (can a node be reached from the entry point?),
(ii)~dead ends (nodes with no outgoing transitions),
(iii)~isolation (must a sensitive tool be preceded by a gate node?),
and (iv)~sanity constraints on router nodes.

These checks are inexpensive: they can be implemented with standard
graph algorithms (BFS/DFS) and run in time linear in $|V| + |E|$.

\subsection{Temporal monitors}
Temporal logics such as linear temporal logic
(LTL)~\citep{pnueli1977temporal} provide a formal language to specify
constraints over execution traces. A common verification strategy is to
compile temporal constraints into automata and reason over traces or
over the product of a system model and the
automaton~\citep{vardi1994automata}.

Agentproof compiles a temporal policy DSL into deterministic finite
automata and supports two complementary evaluation modes:
(i)~\emph{static verification} via a graph~$\times$~DFA product
construction that checks all graph paths without execution, and
(ii)~\emph{runtime monitoring} that evaluates the DFA over a live or
simulated event stream.  The DSL targets the safety fragment of LTL,
deliberately trading full expressiveness for deployment simplicity.

