\subsection{Real-world validation study}
\label{sec:realworld}

The curated corpus (Section~\ref{sec:evaluation}) was designed to contain
representative defects, so its defect rates measure detection capability rather
than prevalence. To estimate \emph{base rates} on independently authored
workflows, and to measure extractor fidelity on code the authors did not write,
we mined public GitHub repositories across all four supported frameworks.

\paragraph{Mining pipeline.}
Using GitHub code search we identified candidate repositories for each
framework, then streamed each repository through clone, extraction, and cleanup
(so disk never accumulates all clones at once). This yielded 930 workflow graphs
from 253 distinct repositories: LangGraph~402, AutoGen~361, CrewAI~114, and
Google~ADK~53. Extraction used the static AST fallback
(\texttt{scripts/ast\_extractor.py}); the runtime extractors of
Section~\ref{sec:graph_model} read the compiled graph object and are expected to
be more faithful, but require per-repository dependency installation that is
infeasible at this scale. Mined workflows are small: a median of 2~non-sentinel
nodes (mean~3.9), predominantly static linear pipelines.

\paragraph{Validation protocol.}
On a stratified 120-workflow sample (40~LangGraph, 20~CrewAI, 35~AutoGen,
25~ADK) we ran two independent LLM-agent passes. First, an agent reconstructed a
\emph{ground-truth} graph from source alone (blind to the extractor output),
following the extractor's node-identifier conventions so the two align; we then
computed node/edge precision--recall and node-kind accuracy against the AST
extraction. Second, for each of the 187 flagged defects, one agent labelled it
\textsc{real}, \textsc{extraction-artifact}, or \textsc{intentional} from
source, and an adversarial verifier independently attempted to overturn each
label, with disagreements recorded as \textsc{arguable}.

\begin{table}[t]
\centering
\small
\caption{AST-extractor fidelity on real GitHub workflows ($n=120$ ground-truth
graphs). Stub-based validation (Section~\ref{sec:accuracy}) reported perfect
precision/recall; on real code fidelity is substantially lower and strongly
framework-dependent---ADK, whose topology is most implicit, is the worst case.}
\label{tab:realworld_fidelity}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{Framework} & \textbf{$n$} & \textbf{Edge P} & \textbf{Edge R} & \textbf{Kind acc} \\
\midrule
LangGraph & 40  & 0.956 & 0.682 & 0.647 \\
CrewAI    & 20  & 0.700 & 0.692 & 0.929 \\
AutoGen   & 35  & 0.957 & 0.770 & 0.900 \\
ADK       & 25  & 0.403 & 0.367 & 0.943 \\
\midrule
\textbf{Overall} & \textbf{120} & \textbf{0.798} & \textbf{0.644} & \textbf{0.829} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Extractor fidelity.}
Table~\ref{tab:realworld_fidelity} reports the result. Node detection stays
strong (overall precision~0.91, recall~0.85), but edge recall falls to 0.644 and
node-kind accuracy to 0.829---far below the perfect scores on programmatic
stubs---and edge recall ranges from 0.37 (ADK) to 0.77 (AutoGen). The frameworks
that encode topology \emph{implicitly} (ADK's nested \texttt{SequentialAgent}/
\texttt{ParallelAgent} composition, AutoGen's participant lists with dynamic
speaker selection) are hardest: the extractor must guess an edge set the source
never states explicitly, and for ADK it is wrong more often than right. On
LangGraph, whose edges are explicit, the dominant error is instead node-kind
(\texttt{tool}$\rightarrow$\texttt{llm}: nodes that invoke tools inside their
body without declared bindings).

\begin{table}[t]
\centering
\small
\caption{Adversarial triage of 187 defects flagged on real workflows. Every
structural flag is an extraction artifact; the human-gate policy flag is
predominantly an intentional design choice.}
\label{tab:realworld_triage}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{Check} & \textbf{Real} & \textbf{Artifact} & \textbf{Intentional} & \textbf{Arguable} \\
\midrule
exit\_reachability     & 0 & 14 & 0  & 0 \\
reverse\_reachability  & 0 & 25 & 0  & 0 \\
dead\_ends             & 0 & 24 & 0  & 1 \\
router\_shape          & 0 & 3  & 0  & 0 \\
tool\_declarations     & 0 & 5  & 0  & 0 \\
human\_presence        & 2 & 9  & 93 & 11 \\
\midrule
\textbf{Total}         & \textbf{2} & \textbf{80} & \textbf{93} & \textbf{12} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Defect prevalence.}
Table~\ref{tab:realworld_triage} gives the result: of 187 flagged defects, only
2 (1.1\%) are genuine bugs---both missing human gates---while 43\% are extraction
artifacts and 50\% are intentional design choices. \emph{Every one of the 71
structural flags is an extraction artifact}, and structural flags are entirely a
LangGraph phenomenon: in the raw study they fire on 78\% of LangGraph workflows
but on 0\% of AutoGen, CrewAI, and ADK workflows, because those extractors build
a connected topology by construction and cannot emit a dead end or unreachable
exit. The structural-defect signal is thus an artifact of the one extractor that
recovers real edges, not a property of the workflows. The \texttt{human\_presence}
policy check fires on 91\% of workflows, but on the validated sample only 2 of
115 firings are genuine (the rest intentional low-risk tasks or artifacts),
reinforcing the context-dependence argued in Section~\ref{sec:evaluation} and
motivating a risk-aware variant.

\paragraph{Takeaway.}
Topology-level defects are \emph{rare} in publicly available agent workflows,
which are mostly small static pipelines, and the curated-benchmark rates do not
transfer to production prevalence. The practical determinants of static-analysis
utility are therefore extraction fidelity---which varies sharply by framework and
collapses on implicit-topology frameworks like ADK---and policy applicability,
rather than the check algorithms themselves. We read this as redirecting the
research agenda toward higher-fidelity extraction (sandboxed runtime extraction;
AST/runtime cross-validation) and risk-aware policy specification, and toward
giving the static machinery a defect-independent job---pruning provably-inert
runtime monitors---rather than cataloguing topology checks whose real-world
firing is dominated by extraction noise.
