\section{Limitations and future work}
\label{sec:limitations}

\paragraph{Workflow structure vs.\ LLM semantics.}
The proposed approach verifies properties of the workflow \emph{topology} and
the event stream emitted by execution. It does not prove semantic
properties of LLM outputs (e.g., factuality or intent), which depend
on non-deterministic model behavior and prompt/context choices.

\paragraph{Extractor heuristics.}
Frameworks evolve quickly and expose different internal representations
across versions. The extractors employ heuristics (e.g., recognizing
human review nodes by naming conventions) that may not generalize to
all codebases. The current validation uses 27 stub-based test cases
(Section~\ref{sec:accuracy}); the gap between stub-based validation and
the diversity of real-world workflow definitions is not yet measured.
The AST-based extractor (\texttt{scripts/\allowbreak ast\_extractor.py})
provides an independent extraction path that could serve as a
cross-validation tool to strengthen confidence in extraction fidelity.

\paragraph{Runtime graph mutation (T3).}
Some frameworks allow dynamic modification of the workflow graph during
execution. If the post-mutation graph is not re-verified, static
guarantees no longer hold. Integrating re-verification hooks into
framework event systems is a natural extension.

\paragraph{Temporal DSL expressiveness.}
The DSL covers seven expression forms targeting the safety fragment of
LTL but does not support full LTL, nested temporal operators beyond
two levels, past-time modalities, or real-time bounds beyond step
counting. All 15 evaluation policies fit within this fragment; however,
these policies were authored by the same team that designed the DSL,
introducing a self-selection bias: requirements that did not fit the
grammar were unlikely to be proposed. The policies were inspired by
external standards (GDPR~\citep{gdpr_article22},
SOC~2~\citep{aicpa_soc2}, OWASP LLM Top~10~\citep{owasp_llm_top10}),
which provides partial mitigation, but a systematic survey of safety
requirements from independent production deployments is needed to
validate DSL sufficiency. Extending to richer temporal fragments
(e.g., past-time LTL, real-time constraints) is future work.

\paragraph{False positive management.}
Some flagged defects may be intentional design choices (e.g., dead-end
nodes used as intentional error-halting states, or workflows that
intentionally omit human gates for low-risk tasks). Future work
includes annotation-based suppression (e.g.,
\texttt{\# agentproof: ignore dead-end} comments), severity tiers,
and configurable verification profiles. The library already supports
a \texttt{suppressions} parameter for node-level exclusions.

\paragraph{Static temporal conservatism.}
The graph $\times$ DFA product construction explores all topological
paths, including paths that may be infeasible at runtime due to router
conditions or LLM decision logic.  This makes the analysis sound
(no false negatives) but potentially conservative (false positives).
Incorporating edge conditions into the product construction to prune
infeasible paths is a direction for future work.

\paragraph{Corpus scale.}
The evaluation uses 18 author-constructed workflows.
\emph{Threat to validity:} because the corpus was designed to include
representative defects and anti-patterns, the observed defect rates
reflect detection capability on a targeted benchmark, not prevalence in
production systems. Generalizing these rates requires validation on
independently authored workflows.
Agentproof provides a GitHub mining pipeline
(\texttt{scripts/\allowbreak scrape\_workflows.py}) and an AST-based
fallback extractor (\texttt{scripts/\allowbreak ast\_extractor.py}) as
ready infrastructure for building larger real-world corpora. A
large-scale study across hundreds of open-source repositories is needed
to establish base rates and to provide a reusable benchmark for the
community.

