\section{Conclusion}
\label{sec:conclusion}

Agent frameworks encode behavior as explicit workflow graphs, and
\textsc{Agentproof} turns that structure into an analyzable,
cross-framework graph model with six structural checks, witness traces,
and a safety-fragment-LTL policy DSL evaluated both statically (via a
graph\,$\times$\,DFA product) and at runtime. Used as a measurement
instrument, it enabled the first base-rate study of topology defects in
publicly available agent workflows---930 mined across all four
frameworks---and the finding is deflationary but genuine: real workflows
are mostly small static pipelines, and only 2 of 187 flagged defects are
real bugs, with the rest extraction artifacts or intentional design
choices. We therefore retire the earlier framing in
which curated-benchmark defect rates stood in for prevalence; the
18-workflow corpus remains a controlled test that the checks fire when
defects are present, not evidence that they are common.

The contribution survives the rarity of defects and is threefold. First,
a measurement instrument and the first real-world base rates for
agent-workflow topology defects---a number nobody had put down. Second, a
quantified diagnosis the field can act on directly: extraction fidelity,
not the check algorithms, is the binding constraint (edge recall 0.64 on
real code, collapsing to 0.37 on implicit-topology frameworks like ADK,
with every triaged structural flag an extraction artifact). Third, a
defect-independent use for the static machinery: the
graph\,$\times$\,DFA product proves which runtime monitors can be soundly
pruned, making pre-deployment analysis a monitor-selection layer within a
static-plus-runtime safety stack rather than a standalone bug detector.
We read the negative prevalence result not as a null result but as a
redirection---toward higher-fidelity extraction and risk-aware policy
specification, the direction we pursue in follow-up work on runtime
shields.

\paragraph{Artifact availability.}
The Agentproof tool, curated corpus, temporal policies, and
all evaluation scripts are available at
\url{https://github.com/NordicAgents/AgentProof} under the MIT
license. A reproducibility script
(\texttt{scripts/reproduce\_all.sh}) runs the complete evaluation
pipeline.

