\pdfoutput=1
\documentclass[11pt]{article}

\usepackage{arxiv}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{microtype}
\usepackage[numbers]{natbib}
\usepackage{xcolor}
\usepackage{url}
\usepackage{hyperref}

\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}

\title{Agentproof: Static Verification of Agent Workflow Graphs}

% Placeholder author block (fill in later).
\author{
  First Last \\
  Affiliation \\
  \texttt{email@example.com}
  \and
  Second Author \\
  Affiliation \\
  \texttt{email@example.com}
}

\date{}

\begin{document}
\maketitle

\begin{abstract}
Agent frameworks increasingly encode tool-using behavior as explicit
workflow graphs, yet safety enforcement remains a runtime concern.
We observe that these frameworks expose analyzable graph structure
through their APIs, enabling pre-deployment static verification of
safety properties that runtime guardrails can only check reactively.

We present \textsc{Agentproof}, a system that automatically extracts a
unified abstract graph model from four major agent frameworks
(LangGraph, CrewAI, AutoGen, Google ADK), applies five structural
checks with witness trace generation, and evaluates temporal safety
policies via a DSL compiled to deterministic finite automata. Unlike
general-purpose model checkers, Agentproof requires no manual modeling.

In a study of 18 curated workflows, we find that 27\% contain
structural defects (dead-end nodes, unreachable exits) and 55\% violate
a human-gate policy---distinct categories that prior work conflates.
All 15 temporal policies we define fit within our seven-form DSL
fragment, and verification completes in sub-second time for graphs up
to 5{,}000 nodes. Our corpus is author-constructed and moderate in size; defect rates
should not be generalized without validation on larger real-world
corpora. Nonetheless, static graph verification complements runtime
guardrails by catching topology-level defects that runtime tools miss
unless the offending path is exercised.
\end{abstract}

\input{sections/01_introduction}
\input{sections/02a_threat_model}
\input{sections/02_background}
\input{sections/03_system_overview}
\input{sections/04_graph_model_and_extraction}
\input{sections/05_verification_methods}
\input{sections/06_evaluation}
\input{sections/09_related_work}
\input{sections/07_limitations_and_future_work}
\input{sections/08_conclusion}

\bibliographystyle{plainnat}
\bibliography{references}

\appendix
\input{appendix_proofs}
\input{appendix_grammar}

\end{document}

